---
title: "Covid91 vaccine study Final2020"
author: "Brooke De La Torre"
date: "10/22/20"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
OldData <-read.csv(file="covid91.csv",header=TRUE)
#recode aids to say covid and then remove the annoying "X" from the dataframe
FinalData <- OldData %>% mutate(infected = recode(infected, aids = "covid", none = "OK" )) %>% select(-one_of("X"))
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->
  
A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine.
  
$H_0$:The null hypothesis is that among the target population, the vaccine has no effect on whether one contracts the virus or not. 
  
$A_0$:The Alternative hypothesis is that among the target population, the vaccine has an effect on whether one contracts the virus or not.
   
# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->
Every question is a cat~cat type.   
  
  
# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Analysis for Gals
  
## Descriptive results

```{r}
DataF = filter(FinalData,sex=="gal")
barchartGC(~infected + treatment,data=DataF)
barchartGC(~infected + treatment,data=DataF, type="percent",main = "Gals")

```
  
Among the females that contracted aids twice as many who took the placebo. Therefore it appears that the vaccine does have a positive effect.

```{r}
table1 <- xtabs(~infected + treatment, data=DataF)
rowPerc(table1)
colPerc(table1)
```

According to the data only 33.11% of infected were Gals who took the drug where as 66.89% of the placebo takers contracted Covid19. This means that double of the females who took the drug were infected. Therefore the drug seems to be positively working to fight Covid19.

## Inferential results

```{r}
fisher.test(table1)
```

The P-value of the Gals results is less than 0.05 so we would reject the null hypothesis.

## Data analysis for Guys

## Descriptive results

```{r}
DataG = filter(FinalData,sex=="guy")
barchartGC(~infected + treatment,data=DataG)
barchartGC(~infected + treatment,data=DataG, type="percent",main = "Guys")

```
   
Among the Males that contracted aids a third as many took the drug than the placebo. Therefore it appears that the vaccine does have a negative effect resulting into higher numbers of infected.

```{r}
table2 <- xtabs(~infected + treatment, data=DataG)
rowPerc(table2)
colPerc(table2)
```

According to the data 58.16% of infected were Guys who took the drug where as 41.87% of the placebo takers contracted Covid19. This means that a third of the males who took the placebo were infected compared to those who took the drug. Therefore the drug seems to be negatively working to fight Covid19.

## Inferential results

```{r}
fisher.test(table2)
```

The P-value of the Guys results is less than 0.05 so we would reject the null hypothesis.

## Data analysis for Gays
   
## Descriptive results
   
```{r}
DataLGBTQ = filter(FinalData,LGBTQ=="gay")
barchartGC(~infected + treatment,data=DataLGBTQ)
barchartGC(~infected + treatment,data=DataLGBTQ, type="percent",main = "Gays")

```
   
Among the gay participants that contracted aids half as many took the drug than the placebo. Therefore it appears that the vaccine does have a negative effect resulting into higher numbers of infected.
   
```{r}
table3 <- xtabs(~infected + treatment, data=DataLGBTQ)
rowPerc(table3)
colPerc(table3)
```
   
According to the data 70.69% of infected were gays who took the drug where as 29.31% of the placebo takers contracted Covid19. This means that a third of the gays who took the placebo were infected compared to those who took the drug. Therefore the drug seems to be negatively working to fight Covid19.
   
## Inferential results

```{r}
fisher.test(table3)
```
   
The P-value of the Gays results is less than 0.05 so we would reject the null hypothesis.
   
## Data analysis for Drug users

## Descriptive results

```{r}
DataU = filter(FinalData,DrugUser=="yes")
barchartGC(~infected + treatment,data=DataU)
barchartGC(~infected + treatment,data=DataU, type="percent",main = "DrugUser")

```
   
Among the DrugUser participants that contracted aids, a third as many took the placebo. Therefore it appears that the vaccine does have a positive effect.
   
```{r}
table4 <- xtabs(~infected + treatment, data=DataU)
rowPerc(table4)
colPerc(table4)
```
   
According to the data only 33.33% of infected were drugusers who took the drug where as 66.67% of the placebo takers contracted Covid19. This means that double of the people who took the drug were infected. Therefore the drug seems to be positively working to fight Covid19.
   
## Inferential results
  
```{r}
fisher.test(table4)
```

The P-value of the Drugusers results is less than 0.05 so we would reject the null hypothesis.
   
# Discussion and Conclusion
  
<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variables causes the other, or whether they perhaps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
  
## Gals 
  
The P-value for the gals is 2.2e-16 meaning it is less than 0.05, so we reject the null hypothesis that among the target population, the vaccine has no effect on whether one contracts the virus or not. According to the numerical and bar graph data only 33.11% of infected were Gals who took the drug where as 66.89% of the placebo takers contracted Covid19. This means that double of the gals who took the drug were infected. This would suggest that the Drug is working as intended to prevent Covid19 for gals. 
   
## Guys
  
The P-value for the guys is 0.0005801 meaning it is less than 0.05, so we reject the null hypothesis that among the target population, the vaccine has no effect on whether one contracts the virus or not.
According to the data 58.16% of infected were Guys who took the drug where as 41.87% of the placebo takers contracted Covid19. This means that a third of the males who took the placebo were infected compared to those who took the drug. This would suggest that the Drug is causing more cases of Covid19 for Guys.

## Gays

The P-value for the gays is 5.741e-06 meaning it is less than 0.05, so we reject the null hypothesis that among the target population, the vaccine has no effect on whether one contracts the virus or not.
According to the data 70.69% of infected were gays who took the drug where as 29.31% of the placebo takers contracted Covid19. This means that a third of the gays who took the placebo were infected compared to those who took the drug. This would suggest that the drug is causing more cases of Covid19 in Gays.

## DrugUsers
   
The P-value for the DrugUsers is 0.002273 meaning it is less than 0.05, so we reject the null hypothesis that among the target population, the vaccine has no effect on whether one contracts the virus or not.According to the data only 33.33% of infected were Drugusers who took the drug where as 66.67% of the placebo takers contracted Covid19. This means that double of the people who took the drug were infected. This would suggest that the Drug is working as intended to prevent Covid19 for DrugUsers. 
   
## All in relation

All four of the subgroups(gals,guys,gays,drugusers) had a p-value less than 0.05 resulting in rejecting the null hypothesis that among the target population, the vaccine has no effect on whether one contracts the virus or not. Meaning that there is an effect between the target population and that of the vaccine. As seen in the data Gals and Drugusers had a positive correlation with the vaccine that resulted in less infected with Covid19. However as seen in the data Guys and Gays had a negative correlation with the vaccine that resulted in more people infected with covid19. This would suggest that there is a relationship based on gender, sexual orientation, and drug usage, the vaccine will increase or decrease your likelihood of catching Covid19. 